---
title: "Run models in SurvBenchmark"
author:
- name: Yunwei Zhang 20210723
  affiliation: 
  - School of Mathematics and Statistics, The University of Sydney, Australia
  - Charles Perkins Centre, The University of Sydney, Australia
output: rmarkdown::html_vignette
code_folding: "hide"
vignette: >
  %\VignetteIndexEntry{KidneyAllocation}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
  
  
```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    warning = FALSE,
    message = FALSE,
    comment = "#>"
)
```

# Introduction & installation

  
`SurvBenchmark` is an R package to run survival models under the cross validation setting to get the evaluation results for each model, user can also apply further analysis based on these results. 
  
  
This vignette will walk through this process starts from pre-processing the data to run those models.  

* note that there are multiple packages imported in `SurvBenchmark`, and therefore, please make sure you successfully have them installed before using `SurvBenchmark`. Those packages are:  
```
    stats,
    parallel,
    foreach,
    lubridate,
    survival,
    dplyr,
    glmnet,
    rms,
    tidyverse,
    caret,
    pec,
    coefplot,
    survAUC,
    gridExtra,
    ggplot2,
    survminer,
    randomForestSRC,
    ggRandomForests,
    penalized,
    DMwR,
    randomForest,
    riskRegression,
    pROC,
    ROCR,
    cvTools,
    MTLR,
    profmem,
    pseudo,
    survivalROC,
    survival,
    survAUC,
    CoxBoost,
    limma,
    partykit,
    coin,
    compound.Cox,
    GenAlgo,
    survivalsvm,
    rmatio,
    survivalmodels,
    reticulate,
    Matrix,
    keras
```

## installation
```{r}
#install.packages("SurvBenchmark")
library(SurvBenchmark)
```

# An example clinical dataset 

* In this section, we use the veteran dataset from the package "survival"

## Data pre-processing
We load the data and check the column names of the data, in `SurvBenchmark`,it is required the two outcome columns: survival time and survival status are named as **time** and **status** respectively. If your data is not named properly, in this first step, you have to change the column names.   

  
  This veteran data example is ok, no need to change the name. 
```{r}
# veteran data
data("veteran")
colnames(veteran)
```

## Run models in SurvBenchmark and get the evaluated results
  
  First of all, we get the survival formula **fitform_ogl** we would like to use in the model:
```{r}
xnam <- paste(colnames(veteran)[c(1,2,5,6,7,8)], sep="")
form=as.formula(paste("survival::Surv(time, status)~ ", paste(xnam, collapse= "+")))
fitform_ogl=form
fitform_ogl
```

Check the distribution of survival time
```{r}
summary(veteran$time)
hist(veteran$time)
```

Get the time sequence to calculate time-dependent AUC values, here, we pick 15 time points between the 1st and the 3rd quantile of all the time period
```{r}
timess=seq(as.numeric(summary(veteran$time)[2]),as.numeric(summary(veteran$time)[5]),(as.numeric(summary(veteran$time)[5])-as.numeric(summary(veteran$time)[2]))/14)

```

Run Coxph model with AIC backward elimination as the feature selection method.  
  
  
* this example returns a result dataframe with 5 rows, each of them is one fold under a 5-fold cross validation setting. Those columns are the corresponding evaluation measurements.
```{r}
formula1=fitform_ogl
formula2=fitform_ogl
formula3=survival::Surv(time,status)~1
formula4=survival::Surv(time,status)~1
result1=bw_cox1_fun(1,veteran,5,fitform_ogl,formula1,formula2,formula3,formula4,timess)
result1
```

Run this exmple under the parallel setting. 

  * You can also run large datasets under parallel to speed up and to note down the run time and memory used
```{r}
start_time <- Sys.time()
Rprof(tf <- "rprof.log",memory.profiling=TRUE)
cox1 <- pbmcapply::pbmclapply(1:20, bw_cox1_fun,veteran,5, fitform_ogl,formula1,formula2,formula3,formula4,timess, mc.cores = 15)
Rprof(NULL)
mm<-summaryRprof(tf,memory = "both")
mm
#saveRDS(mm,"veteran_bw_cox1m.rds")
#saveRDS(cox1,"veteran_bw_cox1.rds")
#cox1<-readRDS("veteran_bw_cox1.rds")
head(cox1)
end_time <- Sys.time()
end_time - start_time
#saveRDS(end_time - start_time,"veteran_bw_cox1t.rds")
```

# An example omics dataset 

## Data pre-processing. 
  
  * In this section, we use the GSE49997_eset dataset from the package "curatedOvarianData"
  
  We load the data and check the column names of the data, in `SurvBenchmark`,it is required the two outcome columns: survival time and survival status are named as **time** and **status** respectively. If your data is not named properly, in this first step, you have to change the column names.   
    Also, to apply Differential Expression (DE) analysis to get DE genes (similarly, for genetic algorithm feature selection), we require the binarised outcome to be named as **os_class**, and you might need to create that column (see this example below).
    
```{r}
#load the data
data("GSE49997_eset")
expmatrix2=Biobase::exprs(GSE49997_eset)
dim(expmatrix2) #check the dimension of the data

expmatrix2_1=t(expmatrix2)
cancerdt2=cbind.data.frame(expmatrix2_1,GSE49997_eset$vital_status,GSE49997_eset$days_to_death)
colnames(cancerdt2)[16049:16050]=c("status","time") #change the name here
cancerdt2$status=as.vector(ifelse(cancerdt2$status=="living",0,1))

cancerdt2_1=cancerdt2[complete.cases(cancerdt2),] # get rid of missing values
dim(cancerdt2_1) 
#check the time and status
table(cancerdt2$status)
summary(cancerdt2$time)
hist(cancerdt2_1$time)
#create the binary outcome column
cancerdt2_1$os_class=as.vector(ifelse(cancerdt2_1$status==1 & 
                                        cancerdt2_1$time <2*365, "poor", 
                                      ifelse(cancerdt2_1$status==0 & 
                                               cancerdt2_1$time >2*365, "good", "not")))

#get the survival formulas
fitform_ogl=survival::Surv(time,status)~.
formula1=fitform_ogl
formula2=fitform_ogl
formula3=survival::Surv(time,status)~1
formula4=survival::Surv(time,status)~1
form1=as.formula(~.)
timess=seq(as.numeric(summary(cancerdt2_1$time)[2]),as.numeric(summary(cancerdt2_1$time)[5]),(as.numeric(summary(cancerdt2_1$time)[5])-as.numeric(summary(cancerdt2_1$time)[2]))/14)

```

## Run the MTLR model with DE feature selection and get the evaluated results
```{r}
# # parallel
# start_time <- Sys.time()
# Rprof(tf <- "rprof.log",memory.profiling=TRUE)
# result2 <- pbmcapply::pbmclapply(1, mtlr_fun2,cancerdt2_1,5, 16047,1000,timess, mc.cores = 15)
# Rprof(NULL)
# mm<-summaryRprof(tf,memory = "both")
# mm
# end_time <- Sys.time()
# time_diff=end_time - start_time
# colnames(result2[[1]])=c("hc","bc","unoc","ghc","br1","br2","br3","br4","br5","br6","a1","a2","a3","a4","a5","a6","a7","a8","a9","a10","a11","a12","a13","a14","a15","a")
# result2=result2[[1]]
# save(mm,time_diff,result2,file="omics_result.rda")

load("omics_result.rda")
#memory
mm
#time
time_diff
#other evaluation metrics
result2
```

# Session Info

```{r}
sessionInfo("SurvBenchmark")
```
